---
name: Santiago Gerardo
surname: Medina
position: "bioinformatician"
address: "1000 E. 50th Street, Kansas City, MO 64110, USA"
phone: +1 816 267 2239
email: "smedina@stowers.org"
profilepic: "./photo.jpg"
twitter: mariecurie
github: mariecurie
date: "`r format(Sys.time(), '%B %Y')`"
output: 
  vitae::latexcv:
    theme: classic
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
library(tibble)
```

# Personal Details

```{r}
tribble(
   ~ question, ~ answer,
   "Date of Birth", "June 17, 1995",
   "Place of Birth", "Guanajuato, Mexico (Mexican nationality)",
   "Marital Status", "Single"
 ) %>%
   brief_entries(question, answer)
```

# Education

```{r}
library(tibble)
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Bachelor in Genomic Sciences", "2013-17", "Universidad Nacional Autónoma de México", "Léon, Mexico"
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```

**2018 - (current):** *I work as a bioinformatician in [Ariel Bazzini's lab](https://research.stowers.org/bazzinilab/) at the Stowers Institute for Medical Research.*

I collaborate with experimentalists by providing help in analyzing omics-data to study how genes are regulated and how that regulation can impact development or affect human diseases. Also, I have the opportunity to carry my own research projects developing statistical models to understand post-transcriptional gene regulation.

# Research Experience

```{r}
tribble(
  ~Year, ~Type, ~Desc, ~Where,
  2014, "Conservation of microRNAs targets in homologous  genes ", "undergraduate summer research, National Laboratory of Genomics for Biodiversity, Mexico, supervised by Dr. Cei Abreu |", "Using microarrays and RNA-seq data I evaluated if the microRNA’s target sites of homologous genes were conserved and also the conservation of expression pattern due to microRNA effect.",
  
  2015, "A pipeline for peak calling and motif discovery in ChIP-nexus data", "Stowers summer scholar, Stowers Institute, MO USA, supervised by Dr. Julia Zeitlinger |", "I benchmarked different tools for ChiP-seq data analysis with the purpose of developing a pipeline to analyze ChiP-nexus data.",
  
  2016, "Evolution of NBS domains in plant resistance genes", "undergraduate research, UNAM-Leon, supervised by Dr. Julio Vega Arreguin |", "I identified NBS domains of immune resistance genes in the genomes of 30 plant species with BLAST searches. Then I looked at the evolution of these domains with phylogenetic approaches.",
  
  2017, "Tissue-specific codon effect on gene expression", "undergraduate thesis, Stowers Institute, supervised by Dr. Ariel Bazzini |", "I employed Bayesian hierarchical GLMs to estimate the effect of each codon on gene expression across human tissues. For this, I analyzed hundreds of RNA-seq profiles from ENCODE. My analysis showed that the effect of each codon on the mRNA level varies across tissues. This is still ongoing work and now, with a successful collaboration of a graduate student in the lab, we are experimentally validation these results in zebrafish." 
) %>% 
  detailed_entries(
    Year, 
    Type,
    Desc,
    Where
  )
```


# Scholarships

```{r}
tribble(
   ~ scholarship, ~ from, ~ year,
   "Colgate Academic Excellence", "UNAM", "2016",
   "Stowers summer scholar", "Stowers Insititute for Medical Research", "2015"
 ) %>%
   brief_entries(scholarship, year, from)
```

# Workshops

```{r}
tribble(
   ~ nombre, ~ from, ~ year,
   "Bioinformatics Summer School", "Institute of Mathematics UNAM, Querétaro", "2016",
   "Statistical Data Analysis for Genome Scale Biology", "Bressanone-Brixen", "2017"
 ) %>%
   brief_entries(nombre, from, year)
```


# Publications

* Wu, Qiushuang, **Santiago Gerardo Medina**, Gopal Kushawah, Michelle Lynn DeVore, Luciana A. Castellano, Jacqelyn M. Hand, Matthew Wright, and Ariel Alejandro Bazzini. "Translation affects mRNA stability in a codon-dependent manner in human cells." Elife 8 (2019): e45396.

* **Santiago Gerardo Medina**, Gopal Kushawah, Maria Jose Blanco, and Ariel Alejandro Bazzini. "Codon content predicts mRNA stability in vertebrates." *in production and to be available in bioRxiv on March 2020*

***

**Languages**: Fully bilingual (Spanish / English)

**outside the lab**, I enjoy reading, mainly non-fiction, and outdoor activities including mountain biking, hiking, and horseback riding.

# References

+ Dr. Julio
+ Dr. Cei
+ Dr. Bazzini

