---
name: Santiago Gerardo
surname: Medina
position: "bioinformatician (Stowers Institute for Medical Research)"
address: "1000 E. 50th Street, Kansas City, MO 64110, USA"
phone: +1 816 267 2239
email: "santiagogmm123@gmail.com"
profilepic: "./photo.jpg"
twitter: mariecurie
github: mariecurie
date: "`r format(Sys.time(), '%B %Y')`"
output: 
  vitae::latexcv:
    theme: classic
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
library(tibble)
```

# Personal Details

```{r}
tribble(
   ~ question, ~ answer,
   "Date of Birth", "June 17, 1995 (gender: male)",
   "Place of Birth", "Guanajuato, Mexico (Mexican nationality)",
   "Marital Status", "Single"
 ) %>%
   brief_entries(question, answer)
```

# Education

```{r}
library(tibble)
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Bachelor in Agro-genomic Sciences", "2013-17", "Escuela Nacional de Estudios Superiores Unidad León, UNAM", ""
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```


# Research Experience

```{r}
tribble(
  ~Year, ~Type, ~Desc, ~Where,
  2018, "Codon Optimality", "Bioinformatician (current position), Stowers Institute, supervised by Dr. Ariel Bazzini. I have worked on multiple projects in the lab in collaboration with other lab members. Also, I have the opportunity to carry my own research projects developing statistical models to understand post-transcriptional gene regulation.", "" ,
  2017, "Tissue-specific codon effect on gene expression", "undergraduate thesis, Stowers Institute, supervised by Dr. Ariel Bazzini. I employed Bayesian hierarchical models to estimate the effect of each codon on gene expression across human tissues. For this, I analyzed hundreds of RNA-seq profiles from ENCODE. My analysis showed that the effect of each codon on the mRNA level varies across tissues. This is still ongoing work and now, with a successful collaboration of a graduate student in the lab, we are experimentally validation these results in zebrafish.", "" ,
  
  2016, "Evolution of NBS domains in plant resistance genes", "undergraduate research, UNAM-Leon, supervised by Dr. Julio Vega Arreguin. I identified NBS domains of immune resistance genes in the genomes of 30 plant species with BLAST searches. Then I looked at the evolution of these domains with phylogenetic approaches.", "",
  
  2015, "A pipeline for peak calling and motif discovery in ChIP-nexus data", "Stowers summer scholar, Stowers Institute, MO USA, supervised by Dr. Julia Zeitlinger. I benchmarked different tools for ChiP-seq data analysis with the purpose of developing a pipeline to analyze ChiP-nexus data.", "",
  
  2014, "Conservation\nof microRNAs targets in homologous  genes ", "undergraduate summer research, National Laboratory of Genomics for Biodiversity, Mexico, supervised by Dr. Cei Abreu. Using microarrays and RNA-seq data I evaluated if the microRNA’s target sites of homologous genes were conserved and also the conservation of expression pattern due to microRNA effect.", ""

) %>% 
  detailed_entries(
    Year, 
    Type,
    Desc,
    Where
  )
```

# Publications

* Wu, Qiushuang, **Santiago Gerardo Medina**, Gopal Kushawah, Michelle Lynn DeVore, Luciana A. Castellano, Jacqelyn M. Hand, Matthew Wright, and Ariel Alejandro Bazzini. "Translation affects mRNA stability in a codon-dependent manner in human cells." Elife 8 (2019): e45396.

* **Santiago Gerardo Medina**, Gopal Kushawah, Maria Jose Blanco, and Ariel Alejandro Bazzini. "Interplay between codon optimality and 3’ UTR cis-elements controls mRNA stability." *in review*

# Meetings


* **Deciphering mRNA decay programs during embryogenesis.** Santiago Gerardo Medina and Ariel A. Bazzini. (poster) *Biological Data Science. CSHL-2018*
*  **A pipeline for peak calling and motif discovery in ChIP-nexus data.** Santiago Gerardo Medina and Julia Zeitlinger. (poster) *XVI National Congress of Plant Biochemistry and Molecular Biology & IX Symposium Mexico /USA. Querétaro-2016*

# Workshops

```{r}
tribble(
   ~ nombre, ~ from, ~ year,
   "Statistical Data Analysis for Genome Scale Biology", "Bressanone-Brixen", "2017",
   "Bioinformatics Summer School", "Institute of Mathematics UNAM, Querétaro", "2016"
 ) %>%
   brief_entries(nombre, from, year)
```

In 2018 I did an intensive one-week [workshop](https://github.com/santiago1234/taller_de_bioinformatica_2018) with students at my university where I introduced them to scientific Python, data analysis, and data visualization.

# Scholarships

```{r}
tribble(
   ~ scholarship, ~ from, ~ year,
   "Colgate Academic Excellence", "UNAM", "2016",
   "Stowers summer scholar", "Stowers Insititute for Medical Research", "2015"
 ) %>%
   brief_entries(scholarship, year, from)
```

# Skills

```{r}
tribble(
   ~ scholarship, ~ year, ~other,
   "Communication", "technical presentation, data analysis report, scientific publications, data visualization", "",
   "Programing", "R/Bioconductor (advanced), R package development, shiny. Python (Scikit Learn, Pandas, NumPy, SciPy, Keras). Bash scripting, Pipeline development: Snakemake. Software development, git, and github", "",
   "Statistics", "machine learning,  data analysis, generalized linear regression, cluster analysis, principal components analysis, cross validation, generalized linear models, Monte Carlo simulation, Bayesian statistics", NA
 ) %>%
   brief_entries(scholarship, other, year)
```


***

+ **Languages**: Fully bilingual (Spanish / English)


 **outside the lab**, I enjoy reading and outdoor activities including mountain biking, hiking, and horseback riding.

# References

+ Dr. [Julio Vega Arreguin](https://scholar.google.com/citations?user=lQThETwAAAAJ&hl=en)                jvega.arreguin@gmail.com
+ Dr. [Cei Abreu-Goodger](http://ceiabreulab.org/)                 cei.abreu@cinvestav.mx
+ Dr. [Ariel Bazzini](https://research.stowers.org/bazzinilab/index.html)                 arb@stowers.org